### 2023-09-02 my_paper/tables/maineffect-proxy.tex

\begin{table}[H]

\caption{Main Effect of Proxy on Second-Generation's Hispanic Self-identification \label{tab:hispbyproxy}}
\centering
\fontsize{12}{14}\selectfont
\begin{tabular}[c]{>{}lllll}
\toprule
Parents Type & All & Hispanic-Hispanic & Hispanic-White & White-Hispanic\\
\midrule
\textbf{Proxy:} &  &  &  & \\
\hspace{1em}\textbf{Mother} & 0.94 & 0.96 & 0.9 & 0.84\\
\hspace{1em}\textbf{Father} & 0.92 & 0.96 & 0.86 & 0.8\\
\hspace{1em}\textbf{Self} & 0.96 & 0.97 & 0.9 & 0.84\\
\hspace{1em}\textbf{Others} & 0.96 & 0.97 & 0.92 & 0.9\\
\bottomrule
\end{tabular}
\end{table}