### 2023-09-02 my_paper/tables/tab01-SumStat.tex

\begin{table}[H]

\caption{Current Population Survey (CPS) Summary Statistics\label{tab:sumstat1}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{\textbf{Overall}} & \multicolumn{3}{c}{\textbf{By Generation}} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-5}
\textbf{Characteristic} & \makecell[l]{\textbf{All Sample} \\N = 1,131,828} & \makecell[c]{\textbf{First} \\N=119,778} & \makecell[c]{\textbf{Second} \\N=761,450} & \makecell[c]{\textbf{Third} \\N=254,699}\\
\midrule
Female & 0.49 & 0.48 & 0.49 & 0.49\\
Hispanic & 0.91 & 0.96 & 0.94 & 0.82\\
Age & \makecell[c]{8.6 \\(5.1)} & \makecell[c]{11.5 \\(4.3)} & \makecell[c]{8.3 \\(5.0)} & \makecell[c]{7.9 \\(5.0)}\\
\makecell[l]{College Graduate: \\ \hspace{1em}Father} & 0.14 & 0.15 & 0.11 & 0.23\\
\makecell[l]{College Graduate: \\ \hspace{1em}Mother} & 0.14 & 0.15 & 0.11 & 0.22\\
\makecell[l]{Total Family Income \\\hspace{1em}(1999 dollars)} & \makecell[c]{39,882 \\(48,692)} & \makecell[c]{31,927 \\(38,804)} & \makecell[c]{36,726 \\(45,353)} & \makecell[c]{53,000 \\(58,984)}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{*} Mean\ \ (SD)}\\
\end{tabular}}
\begin{tablenotes}
\item[1] The sample includes children ages 17 and below who live in intact families. First-generation Hispanic immigrant children that were born in a Spanish-speaking county. Native-born second-generation Hispanic immigrant children with at least one parent born in a Spanish-speaking country. Finally, native-born third-generation Hispanic immigrant children with native-born parents and at least one grand parent born in a Spanish-speaking country.
\item[2] Data source is the 2004-2021 Current Population Survey.
\end{tablenotes}
\end{threeparttable}
\end{table}