### 2023-09-02 my_paper/tables/tab01a-SumStatOverAll.tex

\begin{table}

\caption{CPS Summary Statistics with Skin IAT Scores \label{tab:sumstat1}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lc}
\toprule
Characteristic & All Sample 
N = 1,131,828\\
\midrule
Female & 0.49\\
Hispanic & 0.91\\
Age & 8.6 
(5.1)\\
Fathers with 
 	 College & 0.14\\
Mothers with 
 	 College & 0.14\\
Total Family Income 
 	 (1999 dollars) & 39,882 
(48,692)\\
\bottomrule
\multicolumn{2}{l}{\rule{0pt}{1em}\textsuperscript{1} makecell[l]{\%; Mean \(SD)}}\\
\end{tabular}}
\begin{tablenotes}
\item[1] The samples include children ages 17 and below who live in intact families. First-generation Hispanic immigrant children that were born in a Spanish speaking county. Native born second-generation Hispanic immigrant children with at least one parent born in a Spanish speaking country. Finally, native born third generation Hispanic immigrant children with native born parents and at least one grand parent born in a Spanish speaking country.
\item[2] Data source is the 2004-2021 Current Population Survey.
\end{tablenotes}
\end{threeparttable}
\end{table}