### 2023-09-02 my_paper/tables/tab03-iat_regression_tab.tex

\begin{table}[!h]

\caption{Relationship Between Bias and Self-Reported Hispanic Identity \label{regtab-01}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lccccc}
\toprule
  & \specialcell{(1) \\ $H_i$} & \specialcell{(2) \\ $H_i$} & \specialcell{(3) \\ $H_i$} & \specialcell{(4) \\ $H_i$} & \specialcell{(5) \\ $H_i$}\\
\midrule
Bias & -0.04 & 0.00 & -0.03*** & 0.03 & -0.10**\\
 & (0.03) & (0.05) & (0.01) & (0.03) & (0.04)\\
Female & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
 & (0.00) & (0.00) & (0.00) & (0.00) & (0.00)\\
College Graduate: Mother & -0.05*** & -0.06*** & -0.05*** & -0.05*** & -0.05***\\
 & (0.01) & (0.01) & (0.01) & (0.01) & (0.01)\\
College Graduate: Father & -0.07*** & -0.07*** & -0.06*** & -0.06*** & -0.06***\\
 & (0.00) & (0.00) & (0.01) & (0.01) & (0.01)\\
First Gen & -1.18*** & 0.13*** & 0.36*** & 0.04*** & 0.26***\\
 & (0.02) & (0.01) & (0.03) & (0.01) & (0.03)\\
Second Gen & -1.20*** & 0.11*** & 0.34*** & 0.02*** & 0.24***\\
 & (0.02) & (0.01) & (0.02) & (0.01) & (0.03)\\
\midrule
Observations & 844,481 & 844,481 & 844,481 & 844,481 & 844,481\\
Mean & 0.91 & 0.91 & 0.91 & 0.91 & 0.91\\
State FE &  &  & X & X & \\
Year FE &  & X &  & X & \\
Year $\times$ Region FE &  &  &  &  & X\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a regression the following regression 
                      (\ref{eq:identity_reg_bias}) with different set of
                      fixed effects. I include controls for sex, quartic age, parental education, fraction of Hispanics in a state, and generational, parents' and grandparents' type dummy variables.
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include children ages 17 and below who live in intact families. 
                      First-generation Hispanic immigrant children that were born in a 
                      Spanish-speaking county. Native-born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish-speaking 
                      country. Finally, native-born third-generation Hispanic immigrant children 
                      with native-born parents and at least one grandparent born in a Spanish 
                      speaking country.}
\item[3] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}