### 2023-09-02 my_paper/tables/tab05-iat_regression_tab_byparent_type.tex

\begin{table}[H]

\caption{Relationship Between Bias and Self-Reported Hispanic identity Among Second-Generation Hispanic Immigrants: By Parental Type \label{regtab-byparent-01}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{\makecell[c]{Both Parents \\ from Spanish \\ Speaking Country \\ (HH)}} & \multicolumn{1}{c}{\makecell[c]{Father \\ from Spanish \\ Speaking Country \\ (HW)}} & \multicolumn{1}{c}{\makecell[c]{Mother  \\ from Spanish \\ Speaking Country \\ (WH)}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & \specialcell{(1) \\ $H^2_{ist}$} & \specialcell{(2) \\ $H^2_{ist}$} & \specialcell{(3) \\ $H^2_{ist}$} & \specialcell{(4) \\ $H^2_{ist}$}\\
\midrule
Bias & -0.13** & -0.15** & -0.01 & -0.07\\
 & (0.05) & (0.06) & (0.07) & (0.10)\\
Female & 0.00 & 0.00 & -0.01 & 0.02**\\
 & (0.00) & (0.00) & (0.01) & (0.01)\\
College Graduate: Mother & -0.06*** & -0.04*** & -0.07*** & -0.08***\\
 & (0.01) & (0.01) & (0.01) & (0.02)\\
College Graduate: Father & -0.08*** & -0.04*** & -0.10*** & -0.11***\\
 & (0.01) & (0.01) & (0.02) & (0.01)\\
\midrule
Observations & 560,100 & 405,116 & 88,421 & 66,563\\
Year $\times$ Region FE & X & X & X & X\\
Mean & 0.94 & 0.96 & 0.9 & 0.83\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a heterogeneous effect of regression (\ref{eq:identity_reg_bias}) by 
                      type of parents with region × year fixed effects. 
                      I include controls for sex, quartic age, fraction of Hispanics in a state, and parental education.
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include second-generation Hispanic children ages 17 and below who live in intact families. 
                      Native-born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish-speaking 
                      country.}
\item[3] \footnotesize{Column (1) includes the results to regression (\ref{eq:identity_reg_bias}) on all second-generation immigrants, 
                                        column (2) includes the results to regression (\ref{eq:identity_reg_bias}) on second-generation immigrants that who has a father and mother that were born in a Spanish-speaking country (HH),
                                        column (3) includes the results to regression (\ref{eq:identity_reg_bias}) on second-generation immigrants that who has a father that was born in a Spanish-speaking country and a native-born mother (HW), and
                                        column (4) includes the results to regression (\ref{eq:identity_reg_bias}) on second-generation immigrants that who has a native-born father and a mother that was born in a Spanish-speaking country (WH).}
\item[4] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}