### 2023-09-02 my_paper/tables/tab06-regression_tab_byproxy_skin.tex

\begin{table}[!h]

\caption{Relationship Between Bias and Self-Reported Hispanic Identity: By Proxy Respondent\label{regtab-proxy-01}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Proxy Respondent} \\
\cmidrule(l{3pt}r{3pt}){2-7}
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{White Mother} & \multicolumn{1}{c}{Hispanic Mother} & \multicolumn{1}{c}{White Father} & \multicolumn{1}{c}{Hispanic Father} & \multicolumn{1}{c}{Self} & \multicolumn{1}{c}{Other} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7}
  & \specialcell{(1) \\ $H_{ist}$} & \specialcell{(2) \\ $H_{ist}$} & \specialcell{(3) \\ $H_{ist}$} & \specialcell{(4) \\ $H_{ist}$} & \specialcell{(5) \\ $H_{ist}$} & \specialcell{(6) \\ $H_{ist}$}\\
\midrule
Prejudice Measure & 0.01 & -0.06 & -0.13 & -0.07*** & -0.07 & -0.17**\\
 & (0.11) & (0.03) & (0.11) & (0.02) & (0.05) & (0.07)\\
Female & -0.01 & 0.00 & 0.01 & 0.00*** & 0.00 & 0.00\\
 & (0.01) & (0.00) & (0.01) & (0.00) & (0.01) & (0.00)\\
College Graduate: Mother & -0.02 & -0.03*** & -0.04*** & -0.03*** & -0.03* & -0.03***\\
 & (0.02) & (0.01) & (0.02) & (0.01) & (0.02) & \vphantom{1} (0.01)\\
College Graduate: Father & -0.09*** & -0.05*** & -0.01 & -0.03*** & -0.10*** & -0.06***\\
 & (0.02) & (0.01) & (0.02) & (0.01) & (0.02) & (0.01)\\
First Gen & -0.10* & 0.07*** & 0.03 & 0.09*** & 0.13*** & 0.08***\\
 & (0.06) & (0.01) & (0.05) & (0.02) & (0.02) & (0.02)\\
Second Gen & 0.05* & 0.04*** & 0.06*** & 0.05*** & 0.12*** & 0.07***\\
 & (0.03) & (0.01) & (0.02) & (0.01) & (0.02) & (0.02)\\
\midrule
Observations & 64,052 & 380,625 & 47,233 & 267,690 & 10,472 & 69,301\\
Year $\times$ Region FE & X & X & X & X & X & X\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a heterogeneous effect of regression (\ref{eq:identity_reg_bias}) by 
                      the proxy household respondent with region × year fixed effects. 
                      I include controls for sex, quartic age, fraction of Hispanics in a state, parental education.
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include children ages 17 and below who live in intact families. 
                      First-generation Hispanic immigrant children that were born in a 
                      Spanish-speaking county. Native-born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish-speaking 
                      country. Finally, native-born third-generation Hispanic immigrant children 
                      with native-born parents and at least one grandparent born in a Spanish 
                      speaking country.}
\item[3] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}