### 2023-09-02 my_paper/tables/tab16-iat_regression_tab_byparent_type_county.tex

\begin{table}[!h]

\caption{Hispanic identity and Skin Tone Implicit Bias Among Second Generation Hispanic Immigrants: By Parental Type\label{regtab-byparent-03}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$}\\
\midrule
Implicit Skin Tone Bias & \num{-0.04}* & \num{-0.04} & \num{-0.03} & \num{-0.05}\\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.07}) & (\num{0.10})\\
Female & \num{0.00} & \num{-0.01} & \num{-0.01} & \num{0.02}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.02})\\
College Graduate: Mother & \num{-0.05}*** & \num{-0.04}*** & \num{-0.04}*** & \num{-0.07}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02})\\
College Graduate: Father & \num{-0.08}*** & \num{-0.04}*** & \num{-0.10}*** & \num{-0.11}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.02})\\
\midrule
N & \num{283872} & \num{210304} & \num{40957} & \num{32611}\\
Region FE & X & X & X & X\\
Year FE & X & X & X & X\\
Year-Region FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}