### 2023-09-02 my_paper/tables/tab18-observations-byproxy-byidentity-mom.tex

\begin{table}[!h]

\caption{Tabulation of Identity by Parents Type and Child's Identity: Mother as Proxy Respondent\label{tab:proxyidentitymom}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{HH} & \multicolumn{4}{c}{HW} & \multicolumn{4}{c}{WH} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13}
\multicolumn{1}{c}{1} & \multicolumn{2}{c}{NonHispanic Child} & \multicolumn{2}{c}{Hispanic Child} & \multicolumn{2}{c}{NonHispanic Child} & \multicolumn{2}{c}{Hispanic Child} & \multicolumn{2}{c}{NonHispanic Child} & \multicolumn{2}{c}{Hispanic Child} & \multicolumn{1}{c}{All} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-14}
Hispanic Mother & N & Percent & N & Percent & N & Percent & N & Percent & N & Percent & N & Percent & Percent\\
\midrule
NonHispanic Mom & 5696 & \num{1.36} & 2244 & \num{0.53} & 7583 & \num{1.80} & 22289 & \num{5.30} & 4129 & \num{0.98} & 998 & \num{0.24} & \num{10.22}\\
Hispanic Mom & 4440 & \num{1.06} & 274034 & \num{65.21} & 980 & \num{0.23} & 56302 & \num{13.40} & 3147 & \num{0.75} & 38361 & \num{9.13} & \num{89.78}\\
All & 10136 & \num{2.41} & 276278 & \num{65.75} & 8563 & \num{2.04} & 78591 & \num{18.70} & 7276 & \num{1.73} & 39359 & \num{9.37} & \num{100.00}\\
\bottomrule
\end{tabular}}
\end{table}