### 2023-09-02 my_paper/tables/tab19-observations-byproxy-byidentity-dad.tex

\begin{table}[!h]

\caption{Tabulation of Identity by Parents Type and Child's Identity: Father as Proxy Respondent\label{tab:proxyidentitydad}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{HH} & \multicolumn{4}{c}{HW} & \multicolumn{4}{c}{WH} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13}
\multicolumn{1}{c}{1} & \multicolumn{2}{c}{NonHispanic Child} & \multicolumn{2}{c}{Hispanic Child} & \multicolumn{2}{c}{NonHispanic Child} & \multicolumn{2}{c}{Hispanic Child} & \multicolumn{2}{c}{NonHispanic Child} & \multicolumn{2}{c}{Hispanic Child} & \multicolumn{1}{c}{All} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-14}
Hispanic Father & N & Percent & N & Percent & N & Percent & N & Percent & N & Percent & N & Percent & Percent\\
\midrule
NonHispanic Dad & 3743 & \num{1.72} & 2308 & \num{1.06} & 2244 & \num{1.03} & 1747 & \num{0.80} & 6183 & \num{2.83} & 9224 & \num{4.23} & \num{11.66}\\
Hispanic Dad & 2931 & \num{1.34} & 147103 & \num{67.42} & 1841 & \num{0.84} & 22794 & \num{10.45} & 485 & \num{0.22} & 17601 & \num{8.07} & \num{88.34}\\
All & 6674 & \num{3.06} & 149411 & \num{68.47} & 4085 & \num{1.87} & 24541 & \num{11.25} & 6668 & \num{3.06} & 26825 & \num{12.29} & \num{100.00}\\
\bottomrule
\end{tabular}}
\end{table}