### 2023-09-02 my_paper/tables/tab21-iat_regression_tab_byparent_type_interaction.tex

\begin{longtable}[t]{lcccc}
\caption{Subjective Hispanic identity and Skin Tone Implicit Bias\label{regtab-interaction-by-sample}}\\
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{1}{c}{All Sample} & \multicolumn{1}{c}{March Sup.} & \multicolumn{1}{c}{non-March Sup.} & \multicolumn{1}{c}{\makecell[c]{March Sup \\with income}} \\
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$}\\
\midrule
\endfirsthead
\caption[]{Subjective Hispanic identity and Skin Tone Implicit Bias \textit{(continued)}}\\
\toprule
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$}\\
\midrule
\endhead

\endfoot
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, and parental education.}\\
\endlastfoot
Bias x Hispanic-White Parent & \num{0.07}*** & \num{-0.03} & \num{0.08}*** & \num{-0.03}\\
 & (\num{0.01}) & (\num{0.04}) & (\num{0.02}) & (\num{0.04})\\
Bias x White-Hispanic Parent & \num{0.06}*** & \num{-0.01} & \num{0.07}*** & \num{-0.01}\\
 & (\num{0.02}) & (\num{0.05}) & (\num{0.02}) & (\num{0.05})\\
Implicit Skin Tone Bias & \num{-0.12}*** & \num{-0.15}*** & \num{-0.12}*** & \num{-0.15}***\\
 & (\num{0.01}) & (\num{0.03}) & (\num{0.01}) & (\num{0.03})\\
Hispanic-White Parent & \num{-0.06}*** & \num{-0.05}*** & \num{-0.06}*** & \num{-0.05}***\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & \vphantom{1} (\num{0.00})\\
White-Hispanic Parent & \num{-0.11}*** & \num{-0.10}*** & \num{-0.11}*** & \num{-0.10}***\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
Log Total Family Income &  &  &  & \num{0.00}***\\
 &  &  &  & (\num{0.00})\\
\midrule
N & \num{560100} & \num{82428} & \num{477672} & \num{82428}\\
Mean & \num{0.94} & \num{0.94} & \num{0.94} & \num{0.94}\\
Year-Region FE & X & X & X & X\\*
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{longtable}