### 2023-09-02 my_paper/tables/tab24-iat_regression_tab_byparent_type.tex

\begin{landscape}\begin{table}[!h]

\caption{Subjective Hispanic identity and Skin Tone Implicit Bias Among Adult Second Generation Hispanic Immigrants: By Parental Type (State)\label{regtab-byparent-adults}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccccccccccccc}
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{4}{c}{All Sample} & \multicolumn{4}{c}{March Sup.} & \multicolumn{4}{c}{non-March Sup.} & \multicolumn{4}{c}{\makecell[c]{March Sup. and \\Control for Income}} \\
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9} \cmidrule(l{3pt}r{3pt}){10-10} \cmidrule(l{3pt}r{3pt}){11-11} \cmidrule(l{3pt}r{3pt}){12-12} \cmidrule(l{3pt}r{3pt}){13-13} \cmidrule(l{3pt}r{3pt}){14-14} \cmidrule(l{3pt}r{3pt}){15-15} \cmidrule(l{3pt}r{3pt}){16-16} \cmidrule(l{3pt}r{3pt}){17-17}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$} & \specialcell{(5) \\ $H^2$} & \specialcell{(6) \\ $H^2$} & \specialcell{(7) \\ $H^2$} & \specialcell{(8) \\ $H^2$} & \specialcell{(9) \\ $H^2$} & \specialcell{(10) \\ $H^2$} & \specialcell{(11) \\ $H^2$} & \specialcell{(12) \\ $H^2$} & \specialcell{(13) \\ $H^2$} & \specialcell{(14) \\ $H^2$} & \specialcell{(15) \\ $H^2$} & \specialcell{(16) \\ $H^2$}\\
\midrule
Implicit Skin Tone Bias & \num{0.34}* & \num{-0.01} & \num{0.42} & \num{0.64} & \num{0.46}* & \num{-0.05} & \num{0.60} & \num{1.03}* & \num{0.33}* & \num{-0.01} & \num{0.40} & \num{0.61} & \num{0.46}* & \num{-0.05} & \num{0.60} & \num{1.03}*\\
 & (\num{0.18}) & (\num{0.12}) & (\num{0.35}) & (\num{0.43}) & (\num{0.23}) & (\num{0.16}) & (\num{0.49}) & (\num{0.54}) & (\num{0.18}) & (\num{0.12}) & (\num{0.34}) & (\num{0.43}) & (\num{0.23}) & (\num{0.16}) & (\num{0.49}) & (\num{0.54})\\
Female & \num{0.00} & \num{0.00} & \num{-0.01} & \num{0.04}*** & \num{0.01} & \num{0.00} & \num{-0.02} & \num{0.05}*** & \num{0.00} & \num{0.00} & \num{-0.01} & \num{0.04}*** & \num{0.00} & \num{0.00} & \num{-0.02} & \num{0.05}***\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.00}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02})\\
Log Total Family Income &  &  &  &  &  &  &  &  &  &  &  &  & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00}\\
 &  &  &  &  &  &  &  &  &  &  &  &  & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.01})\\
\midrule
N & \num{237595} & \num{156819} & \num{46175} & \num{34601} & \num{33458} & \num{22350} & \num{6353} & \num{4755} & \num{204137} & \num{134469} & \num{39822} & \num{29846} & \num{33458} & \num{22350} & \num{6353} & \num{4755}\\
Mean & \num{0.92} & \num{0.96} & \num{0.85} & \num{0.8} & \num{0.94} & \num{0.97} & \num{0.89} & \num{0.84} & \num{0.92} & \num{0.96} & \num{0.84} & \num{0.8} & \num{0.94} & \num{0.97} & \num{0.89} & \num{0.84}\\
Year-Region FE & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X\\
\bottomrule
\multicolumn{17}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] I include controls for sex, quartic age, education.
\item[2] This is a sample of second generation adults between the ages of 25 to 54. The sample is restricted to head of households and partners/spouses.
\item[3] Standard errors are clustered on the state level
\end{tablenotes}
\end{threeparttable}}
\end{table}
\end{landscape}