### 2023-09-02 my_paper/tables/tab25-constrained-regression.tex

\begin{table}[!h]

\caption{Constrained Regression on Adult Second Generation Hispanic Immigrants \label{constrained-reg-adults}}
\centering
\resizebox{\linewidth}{!}{
\fontsize{12}{14}\selectfont
\begin{threeparttable}
\begin{tabular}[c]{lllll}
\toprule
Sample & \specialcell{(1) \\ All \\ Sample} & \specialcell{(2) \\ March \\ Sup.} & \specialcell{(3) \\ non-March \\ Sup.} & \specialcell{(4) \\ March Sup. \\ w. Income}\\
\midrule
Bias x Hispanic-White Parent & 0.05 & 0.29 & 0.03 & 0.29\\
 & (0.23) & (0.25) & (0.23) & (0.25)\\
Bias x White-Hispanic & -0.01 & 0.09 & -0.02 & 0.09\\
 & (0.22) & (0.25) & (0.22) & (0.25)\\
Hispanic-White & -0.11 & -0.18** & -0.11 & -0.18**\\
\addlinespace
 & (0.07) & (0.08) & (0.07) & (0.08)\\
White-Hispanic & -0.14* & -0.17* & -0.13* & -0.17*\\
 & (0.07) & (0.09) & (0.07) & (0.09)\\
Implicit Skin Tone Bias & 0.33* & 0.38* & 0.33* & 0.38*\\
 & (0.18) & (0.23) & (0.18) & (0.23)\\
\addlinespace
Female & -0.00*** & -0.00*** & -0.00*** & -0.00***\\
 & (0.00) & (0.00) & (0.00) & \vphantom{2} (0.00)\\
Age & -0.00*** & -0.00*** & -0.00*** & -0.00***\\
 & (0.00) & (0.00) & (0.00) & \vphantom{1} (0.00)\\
educ & -0.00*** & -0.00*** & -0.00*** & -0.00***\\
\addlinespace
 & (0.00) & (0.00) & (0.00) & (0.00)\\
Log Total Family Earnings &  &  &  & -0.00***\\
 &  &  &  & (0.00)\\
Constant & 0.98*** & 1.01*** & 0.97*** & 1.02***\\
 & (0.05) & (0.07) & (0.05) & (0.07)\\
\addlinespace
N & 237595 & 33458 & 204137 & 33458\\
Mean & 0.92 & 0.94 & 0.92 & 0.94\\
\midrule
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] This is a constrained regression where the controls, sex, education, age, and log family income, are forced to be equal.
\item[2] This is a sample of second generation adults between the ages of 25 to 54. The sample is restricted to head of households and partners/spouses.
\end{tablenotes}
\end{threeparttable}}
\end{table}