### 2023-09-02 my_paper/tables/tab26-iat_regression_tab_byparent_type_interaction-adults.tex

\begin{ThreePartTable}
\begin{TableNotes}
\item[1] I include controls for sex, quartic age, and parental education.
\item[2] This is a sample of second generation adults between the ages of 25 to 54. The sample is restricted to head of households and partners/spouses.
\end{TableNotes}
\begin{longtable}[t]{lcccc}
\caption{Subjective Hispanic identity and Skin Tone Implicit Bias: Adult Second Generation Immigrants\label{regtab-interaction-by-sample-adults}}\\
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{1}{c}{All Sample} & \multicolumn{1}{c}{March Sup.} & \multicolumn{1}{c}{non-March Sup.} & \multicolumn{1}{c}{\makecell[c]{March Sup \\with income}} \\
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$}\\
\midrule
\endfirsthead
\caption[]{Subjective Hispanic identity and Skin Tone Implicit Bias: Adult Second Generation Immigrants \textit{(continued)}}\\
\toprule
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$}\\
\midrule
\endhead

\endfoot
\bottomrule
\insertTableNotes
\endlastfoot
Bias x Hispanic-White Parent & \num{0.05} & \num{0.29}* & \num{0.03} & \num{0.29}*\\
 & (\num{0.06}) & (\num{0.18}) & (\num{0.06}) & (\num{0.18})\\
Bias x White-Hispanic Parent & \num{-0.02} & \num{0.09} & \num{-0.03} & \num{0.09}\\
 & (\num{0.07}) & (\num{0.23}) & (\num{0.07}) & (\num{0.23})\\
Implicit Skin Tone Bias & \num{0.33}*** & \num{0.38}*** & \num{0.32}*** & \num{0.38}***\\
 & (\num{0.04}) & (\num{0.14}) & (\num{0.04}) & (\num{0.14})\\
Hispanic-White Parent & \num{-0.11}*** & \num{-0.18}*** & \num{-0.11}*** & \num{-0.18}***\\
 & (\num{0.02}) & (\num{0.05}) & (\num{0.02}) & (\num{0.05})\\
White-Hispanic Parent & \num{-0.13}*** & \num{-0.16}** & \num{-0.13}*** & \num{-0.16}**\\
 & (\num{0.02}) & (\num{0.07}) & (\num{0.02}) & (\num{0.07})\\
Log Total Family Income &  &  &  & \num{0.00}\\
 &  &  &  & (\num{0.00})\\
\midrule
N & \num{237595} & \num{33458} & \num{204137} & \num{33458}\\
Mean & \num{0.92} & \num{0.94} & \num{0.92} & \num{0.94}\\
Year-Region FE & X & X & X & X\\*
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{longtable}
\end{ThreePartTable}