### 2023-09-02 my_paper/tables/tab29-gss_regression_tab_byparent_type.tex

\begin{landscape}\begin{table}[!h]

\caption{Subjective Hispanic identity and GSS Bias Among Second Generation Hispanic Immigrants: By Parental Type (State)\label{regtab-byparent-04}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccccccccccccc}
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{4}{c}{All Sample} & \multicolumn{4}{c}{March Sup.} & \multicolumn{4}{c}{non-March Sup.} & \multicolumn{4}{c}{March Sup. and Control for Income} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13} \cmidrule(l{3pt}r{3pt}){14-17}
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9} \cmidrule(l{3pt}r{3pt}){10-10} \cmidrule(l{3pt}r{3pt}){11-11} \cmidrule(l{3pt}r{3pt}){12-12} \cmidrule(l{3pt}r{3pt}){13-13} \cmidrule(l{3pt}r{3pt}){14-14} \cmidrule(l{3pt}r{3pt}){15-15} \cmidrule(l{3pt}r{3pt}){16-16} \cmidrule(l{3pt}r{3pt}){17-17}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$} & \specialcell{(5) \\ $H^2$} & \specialcell{(6) \\ $H^2$} & \specialcell{(7) \\ $H^2$} & \specialcell{(8) \\ $H^2$} & \specialcell{(9) \\ $H^2$} & \specialcell{(10) \\ $H^2$} & \specialcell{(11) \\ $H^2$} & \specialcell{(12) \\ $H^2$} & \specialcell{(13) \\ $H^2$} & \specialcell{(14) \\ $H^2$} & \specialcell{(15) \\ $H^2$} & \specialcell{(16) \\ $H^2$}\\
\midrule
Bias & \num{0.03} & \num{0.01} & \num{0.12}** & \num{-0.10} & \num{0.02} & \num{-0.01} & \num{0.00} & \num{-0.02} & \num{0.04} & \num{0.01} & \num{0.13}** & \num{-0.10} & \num{0.02} & \num{-0.01} & \num{0.01} & \num{-0.03}\\
 & (\num{0.03}) & (\num{0.03}) & (\num{0.10}) & (\num{0.12}) & (\num{0.03}) & (\num{0.03}) & (\num{0.10}) & (\num{0.20}) & (\num{0.03}) & (\num{0.03}) & (\num{0.10}) & (\num{0.12}) & (\num{0.03}) & (\num{0.03}) & (\num{0.09}) & (\num{0.20})\\
Female & \num{0.00} & \num{0.00} & \num{-0.01}** & \num{0.01}** & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.02}** & \num{0.00} & \num{0.00} & \num{-0.01}** & \num{0.01}** & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.02}**\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.01}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.01}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.02})\\
College Graduate: Mother & \num{-0.08}*** & \num{-0.04}*** & \num{-0.11}*** & \num{-0.11}*** & \num{-0.08}*** & \num{-0.05}*** & \num{-0.09}*** & \num{-0.10}*** & \num{-0.08}*** & \num{-0.04}*** & \num{-0.11}*** & \num{-0.11}*** & \num{-0.07}*** & \num{-0.05}*** & \num{-0.08}*** & \num{-0.09}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02})\\
College Graduate: Father & \num{-0.10}*** & \num{-0.04}*** & \num{-0.16}*** & \num{-0.15}*** & \num{-0.10}*** & \num{-0.04}*** & \num{-0.17}*** & \num{-0.13}*** & \num{-0.10}*** & \num{-0.04}*** & \num{-0.16}*** & \num{-0.16}*** & \num{-0.10}*** & \num{-0.04}*** & \num{-0.16}*** & \num{-0.12}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.03}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.03}) & (\num{0.02})\\
Log Total Family Income &  &  &  &  &  &  &  &  &  &  &  &  & \num{0.00}*** & \num{0.00}* & \num{-0.01}*** & \num{-0.02}***\\
 &  &  &  &  &  &  &  &  &  &  &  &  & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
\midrule
Mean & \num{0.92} & \num{0.96} & \num{0.83} & \num{0.77} & \num{0.93} & \num{0.97} & \num{0.85} & \num{0.8} & \num{0.92} & \num{0.96} & \num{0.82} & \num{0.77} & \num{0.93} & \num{0.97} & \num{0.85} & \num{0.8}\\
N & \num{293112} & \num{209966} & \num{47324} & \num{35822} & \num{41973} & \num{30332} & \num{6675} & \num{4966} & \num{251139} & \num{179634} & \num{40649} & \num{30856} & \num{41973} & \num{30332} & \num{6675} & \num{4966}\\
Year-Region FE & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X\\
\bottomrule
\multicolumn{17}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{17}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{17}{l}{\rule{0pt}{1em}** p $<$ 0.29, * p $<$ 0.1, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}
\end{landscape}