### 2023-09-02 my_paper/tables/tab31-iat_regression_tab_byparent_type-interethnic.tex

\begin{landscape}\begin{table}[!h]

\caption{Subjective Hispanic identity and Skin Tone Implicit Bias Among Second Generation Hispanic Immigrants: By Parental Type (State)\label{regtab-byparent-04}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccccccccc}
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{3}{c}{All Sample} & \multicolumn{3}{c}{March Sup.} & \multicolumn{3}{c}{non-March Sup.} & \multicolumn{3}{c}{March Sup. and Control for Income} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{Inter Ethnic} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{Inter Ethnic} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{Inter Ethnic} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{Inter Ethnic} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9} \cmidrule(l{3pt}r{3pt}){10-10} \cmidrule(l{3pt}r{3pt}){11-11} \cmidrule(l{3pt}r{3pt}){12-12} \cmidrule(l{3pt}r{3pt}){13-13}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$} & \specialcell{(5) \\ $H^2$} & \specialcell{(6) \\ $H^2$} & \specialcell{(7) \\ $H^2$} & \specialcell{(8) \\ $H^2$} & \specialcell{(9) \\ $H^2$} & \specialcell{(10) \\ $H^2$} & \specialcell{(11) \\ $H^2$} & \specialcell{(12) \\ $H^2$}\\
\midrule
Implicit Skin Tone Bias & \num{-0.10}** & \num{-0.15}** & \num{0.06} & \num{-0.15}*** & \num{-0.16}** & \num{-0.09} & \num{-0.10}* & \num{-0.15}** & \num{0.07} & \num{-0.15}*** & \num{-0.16}** & \num{-0.09}\\
 & (\num{0.05}) & (\num{0.07}) & (\num{0.07}) & (\num{0.05}) & (\num{0.06}) & (\num{0.12}) & (\num{0.05}) & (\num{0.07}) & (\num{0.06}) & (\num{0.05}) & (\num{0.06}) & (\num{0.12})\\
Female & \num{0.00} & \num{0.00} & \num{0.01} & \num{0.00} & \num{0.00} & \num{0.01} & \num{0.00} & \num{0.00} & \num{0.01} & \num{0.00} & \num{0.00} & \num{0.01}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.01})\\
College Graduate: Mother & \num{-0.06}*** & \num{-0.04}*** & \num{-0.07}*** & \num{-0.05}*** & \num{-0.03}** & \num{-0.07}*** & \num{-0.06}*** & \num{-0.04}*** & \num{-0.07}*** & \num{-0.05}*** & \num{-0.03}** & \num{-0.07}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
College Graduate: Father & \num{-0.08}*** & \num{-0.04}*** & \num{-0.12}*** & \num{-0.09}*** & \num{-0.05}*** & \num{-0.11}*** & \num{-0.08}*** & \num{-0.04}*** & \num{-0.12}*** & \num{-0.08}*** & \num{-0.05}*** & \num{-0.11}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02})\\
Log Total Family Income &  &  &  &  &  &  &  &  &  & \num{0.00}*** & \num{0.00} & \num{-0.01}***\\
 &  &  &  &  &  &  &  &  &  & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
\midrule
Mean & \num{0.94} & \num{0.96} & \num{0.9} & \num{0.94} & \num{0.97} & \num{0.9} & \num{0.94} & \num{0.96} & \num{0.89} & \num{0.94} & \num{0.97} & \num{0.9}\\
N & \num{560100} & \num{405116} & \num{154984} & \num{82428} & \num{59950} & \num{22478} & \num{477672} & \num{345166} & \num{132506} & \num{82428} & \num{59950} & \num{22478}\\
Year-Region FE & X & X & X & X & X & X & X & X & X & X & X & X\\
\bottomrule
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{13}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}
\end{landscape}