### 2023-09-02 my_paper/tables/tab32-Explicit_regression_tab_byparent_type.tex

\begin{landscape}\begin{table}[H]

\caption{Subjective Hispanic identity and Explicit Bias Among Second Generation Hispanic Immigrants: By Parental Type (State)\label{regtab-byparent-05}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccccccccccccc}
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{4}{c}{All Sample} & \multicolumn{4}{c}{March Sup.} & \multicolumn{4}{c}{non-March Sup.} & \multicolumn{4}{c}{March Sup. and Control for Income} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13} \cmidrule(l{3pt}r{3pt}){14-17}
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9} \cmidrule(l{3pt}r{3pt}){10-10} \cmidrule(l{3pt}r{3pt}){11-11} \cmidrule(l{3pt}r{3pt}){12-12} \cmidrule(l{3pt}r{3pt}){13-13} \cmidrule(l{3pt}r{3pt}){14-14} \cmidrule(l{3pt}r{3pt}){15-15} \cmidrule(l{3pt}r{3pt}){16-16} \cmidrule(l{3pt}r{3pt}){17-17}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$} & \specialcell{(5) \\ $H^2$} & \specialcell{(6) \\ $H^2$} & \specialcell{(7) \\ $H^2$} & \specialcell{(8) \\ $H^2$} & \specialcell{(9) \\ $H^2$} & \specialcell{(10) \\ $H^2$} & \specialcell{(11) \\ $H^2$} & \specialcell{(12) \\ $H^2$} & \specialcell{(13) \\ $H^2$} & \specialcell{(14) \\ $H^2$} & \specialcell{(15) \\ $H^2$} & \specialcell{(16) \\ $H^2$}\\
\midrule
Explicit Bias & \num{-0.05} & \num{-0.09} & \num{0.08} & \num{0.04} & \num{0.02} & \num{-0.06} & \num{0.49}*** & \num{-0.12} & \num{-0.05} & \num{-0.10} & \num{0.05} & \num{0.05} & \num{0.02} & \num{-0.06} & \num{0.49}*** & \num{-0.12}\\
 & (\num{0.06}) & (\num{0.08}) & (\num{0.11}) & (\num{0.19}) & (\num{0.05}) & (\num{0.07}) & (\num{0.11}) & (\num{0.21}) & (\num{0.06}) & (\num{0.08}) & (\num{0.11}) & (\num{0.20}) & (\num{0.05}) & (\num{0.07}) & (\num{0.11}) & (\num{0.21})\\
Female & \num{0.00} & \num{0.00} & \num{-0.02} & \num{0.02} & \num{0.00} & \num{0.00} & \num{-0.01} & \num{0.02} & \num{0.00} & \num{0.00} & \num{-0.02} & \num{0.02} & \num{0.00} & \num{0.00} & \num{-0.01} & \num{0.02}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.02}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) & (\num{0.02}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) & (\num{0.02}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) & (\num{0.02}) & (\num{0.02})\\
College Graduate: Mother & \num{-0.06}*** & \num{-0.03} & \num{-0.04}** & \num{-0.12}*** & \num{-0.05}** & \num{-0.02} & \num{-0.01} & \num{-0.16}*** & \num{-0.06}*** & \num{-0.03} & \num{-0.04}** & \num{-0.12}*** & \num{-0.05}** & \num{-0.02} & \num{0.00} & \num{-0.16}***\\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03}) & (\num{0.03})\\
College Graduate: Father & \num{-0.08}*** & \num{-0.04}*** & \num{-0.17}*** & \num{-0.07}** & \num{-0.09}*** & \num{-0.05}*** & \num{-0.16}*** & \num{-0.07}** & \num{-0.08}*** & \num{-0.04}*** & \num{-0.17}*** & \num{-0.06}** & \num{-0.09}*** & \num{-0.05}*** & \num{-0.15}*** & \num{-0.07}**\\
 & (\num{0.02}) & (\num{0.01}) & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.01}) & (\num{0.05}) & (\num{0.03}) & (\num{0.02}) & (\num{0.01}) & (\num{0.03}) & (\num{0.03}) & (\num{0.03}) & (\num{0.01}) & (\num{0.05}) & (\num{0.03})\\
Log Total Family Income &  &  &  &  &  &  &  &  &  &  &  &  & \num{0.00}*** & \num{0.00}*** & \num{-0.01} & \num{-0.01}**\\
 &  &  &  &  &  &  &  &  &  &  &  &  & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
\midrule
Mean & \num{0.94} & \num{0.96} & \num{0.9} & \num{0.83} & \num{0.94} & \num{0.97} & \num{0.9} & \num{0.85} & \num{0.94} & \num{0.96} & \num{0.89} & \num{0.83} & \num{0.94} & \num{0.97} & \num{0.9} & \num{0.85}\\
N & \num{167005} & \num{122157} & \num{25648} & \num{19200} & \num{24844} & \num{18157} & \num{3846} & \num{2841} & \num{142161} & \num{104000} & \num{21802} & \num{16359} & \num{24844} & \num{18157} & \num{3846} & \num{2841}\\
Year $\times$ Region FE & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X\\
\bottomrule
\multicolumn{17}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] I include controls for sex, quartic age, parental education.
\item[2] Standard errors are clustered on the state level
\end{tablenotes}
\end{threeparttable}}
\end{table}
\end{landscape}