### 2023-09-02 my_paper/tables/tab36-blackiat_regression_tab.tex

\begin{table}[!h]

\caption{Subjective Black Identity and Race Implicit Bias (State) \label{regtab-05}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lccccc}
\toprule
  & \specialcell{(1) \\ $B_i$} & \specialcell{(2) \\ $B_i$} & \specialcell{(3) \\ $B_i$} & \specialcell{(4) \\ $B_i$} & \specialcell{(5) \\ $B_i$}\\
\midrule
Implicit Bias & \num{-0.13} & \num{-0.09} & \num{-0.10} & \num{-0.05} & \num{0.02}\\
 & (\num{0.14}) & (\num{0.10}) & (\num{0.12}) & (\num{0.09}) & (\num{0.08})\\
Female & \num{-0.01} & \num{-0.01} & \num{-0.01} & \num{-0.01} & \num{-0.01}\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
College Graduate: Mother & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00}\\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
College Graduate: Father & \num{-0.03}** & \num{-0.04}** & \num{-0.03}** & \num{-0.04}** & \num{-0.03}**\\
 & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02})\\
\midrule
N & \num{95327} & \num{95327} & \num{95327} & \num{95327} & \num{95327}\\
Region FE &  &  & X & X & \\
Year FE &  & X &  & X & \\
Mean & \num{0.7} & \num{0.7} & \num{0.7} & \num{0.7} & \num{0.7}\\
NA & NA & NA & NA & NA & \vphantom{3} NA\\
NA & NA & NA & NA & NA & \vphantom{2} NA\\
NA & NA & NA & NA & NA & \vphantom{1} NA\\
NA & NA & NA & NA & NA & NA\\
Year-Region FE &  &  &  &  & X\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{3} Data source is the current population survey.}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}