### 2023-09-02 my_paper/tables/tab37-blackiat_regression_tab_byparent_type.tex

\begin{landscape}\begin{table}[!h]

\caption{Subjective Black Identity and Implicit Bias: By Parental Type (State)\label{regtab-byparent-07}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccccccccc}
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{3}{c}{All Sample} & \multicolumn{3}{c}{March Sup.} & \multicolumn{3}{c}{non-March Sup.} & \multicolumn{3}{c}{March Sup. and Control for Income} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{BW} & \multicolumn{1}{c}{WB} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{BW} & \multicolumn{1}{c}{WB} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{BW} & \multicolumn{1}{c}{WB} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{BW} & \multicolumn{1}{c}{WB} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9} \cmidrule(l{3pt}r{3pt}){10-10} \cmidrule(l{3pt}r{3pt}){11-11} \cmidrule(l{3pt}r{3pt}){12-12} \cmidrule(l{3pt}r{3pt}){13-13}
  & \specialcell{(1) \\ $B_i$} & \specialcell{(2) \\ $B_i$} & \specialcell{(3) \\ $B_i$} & \specialcell{(4) \\ $B_i$} & \specialcell{(5) \\ $B_i$} & \specialcell{(6) \\ $B_i$} & \specialcell{(7) \\ $B_i$} & \specialcell{(8) \\ $B_i$} & \specialcell{(9) \\ $B_i$} & \specialcell{(10) \\ $B_i$} & \specialcell{(11) \\ $B_i$} & \specialcell{(12) \\ $B_i$}\\
\midrule
Implicit Bias & \num{0.02} & \num{0.07} & \num{-0.10} & \num{0.03} & \num{0.08} & \num{-0.04} & \num{0.02} & \num{0.07} & \num{-0.11} & \num{0.03} & \num{0.08} & \num{-0.04}\\
 & (\num{0.08}) & (\num{0.09}) & (\num{0.10}) & (\num{0.09}) & (\num{0.10}) & (\num{0.15}) & (\num{0.08}) & (\num{0.09}) & (\num{0.11}) & (\num{0.09}) & (\num{0.10}) & (\num{0.15})\\
Female & \num{-0.01} & \num{-0.01} & \num{-0.02} & \num{-0.02}*** & \num{-0.02} & \num{-0.04} & \num{-0.01} & \num{-0.01} & \num{-0.02} & \num{-0.02}*** & \num{-0.02} & \num{-0.04}\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02})\\
College Graduate: Mother & \num{0.00} & \num{0.00} & \num{-0.02} & \num{0.03} & \num{0.04}* & \num{-0.01} & \num{-0.01} & \num{0.00} & \num{-0.03} & \num{0.03}* & \num{0.04}** & \num{-0.02}\\
 & (\num{0.01}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.04}) & (\num{0.01}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.04})\\
College Graduate: Father & \num{-0.03}** & \num{-0.05}** & \num{0.01} & \num{-0.05}** & \num{-0.06}** & \num{-0.03} & \num{-0.03}** & \num{-0.05}** & \num{0.01} & \num{-0.05}** & \num{-0.06}** & \num{-0.03}\\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.03}) & (\num{0.04}) & (\num{0.01}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.03}) & (\num{0.04})\\
Log Total Family Income &  &  &  &  &  &  &  &  &  & \num{0.00} & \num{-0.01}** & \num{0.00}\\
 &  &  &  &  &  &  &  &  &  & (\num{0.00}) & (\num{0.00}) & (\num{0.01})\\
\midrule
Mean & \num{0.7} & \num{0.7} & \num{0.69} & \num{0.7} & \num{0.7} & \num{0.7} & \num{0.7} & \num{0.7} & \num{0.69} & \num{0.7} & \num{0.7} & \num{0.7}\\
N & \num{95327} & \num{69169} & \num{26158} & \num{12123} & \num{8778} & \num{3345} & \num{83204} & \num{60391} & \num{22813} & \num{12123} & \num{8778} & \num{3345}\\
Year-Region FE & X & X & X & X & X & X & X & X & X & X & X & X\\
\bottomrule
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{13}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}
\end{landscape}