### 2023-09-02 my_paper/tables/tab39-iat_regression_tab_byparent_type-ints.tex

\begin{table}[H]

\caption{Self-Reported Hispanic identity and Bias Among Second Generation Hispanic Immigrants: By Parental Type (Bias Averages Over Three Years Intervals) \label{regtab-byparent-ints}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{HH} & \multicolumn{1}{c}{HW} & \multicolumn{1}{c}{WH} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$}\\
\midrule
Bias & \num{-0.24}** & \num{-0.16} & \num{-0.31}** & \num{-0.48}**\\
 & (\num{0.10}) & (\num{0.13}) & (\num{0.12}) & (\num{0.22})\\
Female & \num{0.00} & \num{0.00} & \num{-0.01} & \num{0.02}**\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.01})\\
College Graduate: Mother & \num{-0.06}*** & \num{-0.04}*** & \num{-0.06}*** & \num{-0.08}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.02})\\
College Graduate: Father & \num{-0.08}*** & \num{-0.04}*** & \num{-0.11}*** & \num{-0.11}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.02}) & (\num{0.01})\\
\midrule
N & \num{560100} & \num{405116} & \num{88421} & \num{66563}\\
Region FE & X & X & X & X\\
Mean & \num{0.94} & \num{0.96} & \num{0.9} & \num{0.83}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a heterogeneous effect of regression (\ref{eq:identity_reg_bias}) by 
                      generation with region fixed effexts. 
                      I include controls for sex, quartic age, parental education.
                      Standard errors are clustered on the state level}
\item[2] \footnotesize{The samples include second-generation Hispanic children ages 17 and below who live in intact families. 
                      Native born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish speaking 
                      country.}
\item[3] \footnotesize{Data source is the 2004-2020 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}