### 2023-09-02 my_paper/tables/tab42-iat-regression-parent-type.tex

\begin{table}[H]

\caption{Relationship Between Parents' Type and Bias Among Second Generation Hispanic Immigrants\label{regtab-skin-iat-parents-type}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{lccc}
\toprule
  & \specialcell{(1) \\ $HH$} & \specialcell{(2) \\ $HW$} & \specialcell{(3) \\ $WH$}\\
\midrule
Bias & \num{-0.05}* & \num{-0.02} & \num{0.06}**\\
 & (\num{0.03}) & (\num{0.02}) & (\num{0.03})\\
College Graduate: Wife & \num{-0.05}*** & \num{0.02}*** & \num{0.03}***\\
 & (\num{0.01}) & (\num{0.00}) & \vphantom{1} (\num{0.01})\\
College Graduate: Husband & \num{-0.08}*** & \num{0.02}*** & \num{0.06}***\\
 & (\num{0.01}) & (\num{0.00}) & (\num{0.01})\\
Wife's Age & \num{0.00}*** & \num{0.00}*** & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & \vphantom{3} (\num{0.00})\\
Husband's Age & \num{0.00} & \num{0.00} & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & \vphantom{2} (\num{0.00})\\
Year Immigrated: Wife & \num{0.00}*** & \num{0.00}*** & \num{0.00}***\\
 & (\num{0.00}) & (\num{0.00}) & \vphantom{1} (\num{0.00})\\
Year Immigrated: Husband & \num{0.00}*** & \num{0.00}*** & \num{0.00}***\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
\midrule
N & \num{560100} & \num{560100} & \num{560100}\\
Year $\times$ Region FE & X & X & X\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{This is the result to estimating (\ref{eq:inter-hw}) as a
                      linear probability model.}
\item[2] \footnotesize{I include controls for partners' sex, age, education, and years since immigrating to the United States.
                      Standard errors are clustered on the state level}
\item[3] \footnotesize{The samples include second-generation Hispanic children ages 17 and below who live in intact families. 
                      Native born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish speaking 
                      country.}
\item[4] \footnotesize{Data source is the 2004-2020 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}
\end{table}