### 2023-09-02 my_paper/tables/tab45-iat_regression_tab_bygrandparent_type.tex

\begin{table}[H]

\caption{Relationship Between Bias and Self-Reported Hispanic identity Among Third-Generation Hispanic Immigrants: By Grandparental Type \label{regtab-bygrandparents}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Number of Hispanic Grandparents} \\
\cmidrule(l{3pt}r{3pt}){2-5}
  & \specialcell{(1) \\ One} & \specialcell{(2) \\ Two} & \specialcell{(3) \\ Three} & \specialcell{(4) \\ Four}\\
\midrule
Bias & -0.04 & 0.03 & 0.19 & -0.14*\\
 & (0.11) & (0.09) & (0.26) & (0.07)\\
Female & -0.01 & 0.00 & -0.01 & 0.00\\
 & (0.01) & (0.01) & (0.01) & (0.01)\\
College Graduate: Mother & -0.11*** & -0.07*** & 0.02 & -0.02\\
 & (0.03) & (0.02) & (0.02) & (0.01)\\
College Graduate: Father & -0.11*** & -0.08*** & 0.02 & -0.03*\\
 & (0.03) & (0.01) & (0.01) & (0.01)\\
\midrule
Observations & 55,051 & 74,100 & 12,194 & 57,646\\
Year $\times$ Region FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of equation (\ref{eq:identity_reg_bias}) restricted to third-generation Hispanic immigrants by 
                      number of Hispanic grandparents with region × year fixed effects. 
                      I include controls for sex, quartic age, fraction of Hispanics in a state, and parental education.
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include third-generation Hispanic children ages 17 and below who live in intact families. 
                      Native-born third-generation Hispanic 
                      immigrant children with at least one grandparent born in a Spanish-speaking 
                      country.}
\item[3] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}