### 2023-09-02 my_paper/tables/tab47-regression-reverse-causality.tex

\begin{table}[!h]

\caption{Effect of Hispanic on Bias: By Generation \label{regtab-reverse-causality}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ All Gens \\ $Bias_{st}$} & \specialcell{(2) \\  First Gen \\ $Bias_{st}$} & \specialcell{(3) \\  Second Gen \\ $Bias_{st}$} & \specialcell{(4) \\  Third Gen \\ $Bias_{st}$}\\
\midrule
Hispanic & \num{0.00} & \num{0.00}** & \num{0.00}* & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & \vphantom{3} (\num{0.00})\\
Female & \num{0.00}* & \num{0.00} & \num{0.00}* & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & \vphantom{2} (\num{0.00})\\
College Graduate: Mother & \num{0.00} & \num{0.01}** & \num{0.00} & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & \vphantom{1} (\num{0.00})\\
College Graduate: Father & \num{0.00} & \num{0.00}* & \num{0.00}** & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
\midrule
N & \num{844481} & \num{85390} & \num{560100} & \num{198991}\\
Mean & \num{0.91} & \num{0.96} & \num{0.94} & \num{0.82}\\
Year $\times$ Region FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{The samples include children ages 17 and below who live in intact families. 
                      First-generation Hispanic immigrant children that were born in a 
                      Spanish speaking county. Native born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish speaking 
                      country. Finally, native born third generation Hispanic immigrant children 
                      with native born parents and at least one grand parent born in a Spanish 
                      speaking country.}
\item[2] \footnotesize{Data source is the 2004-2020 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}