### 2023-09-02 my_paper/tables/tab49-iat_regression_tab_byparent_type-specs.tex

\begin{table}[H]

\caption{Self-Reported Hispanic identity and Bias Among Second Generation Hispanic Immigrants: Both Parents Born in a Spanish Speaking Country \label{regtab-byparent-specifications}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccc}
\toprule
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$} & \specialcell{(5) \\ $H^2$} & \specialcell{(6) \\ $H^2$}\\
\midrule
Bias & \num{-0.01} & \num{-0.02} & \num{-0.03} & \num{-0.04} & \num{-0.02} & \num{-0.15}**\\
 & (\num{0.03}) & (\num{0.06}) & (\num{0.02}) & (\num{0.03}) & (\num{0.02}) & (\num{0.07})\\
Female & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
College Graduate: Mother & \num{-0.05}*** & \num{-0.05}*** & \num{-0.03}*** & \num{-0.03}*** & \num{-0.03}*** & \num{-0.04}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & \vphantom{1} (\num{0.01})\\
College Graduate: Father & \num{-0.05}*** & \num{-0.05}*** & \num{-0.04}*** & \num{-0.04}*** & \num{-0.04}*** & \num{-0.04}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
\midrule
N & \num{405116} & \num{405116} & \num{405116} & \num{405116} & \num{405116} & \num{405116}\\
Linear Trend & No & No & No & No & Yes & No\\
Mean & 0.96 & 0.96 & 0.96 & 0.96 & 0.96 & 0.96\\
Year FE &  & X &  & X &  & \\
State FE &  &  & X & X &  & \\
Year $\times$ Region FE &  &  &  &  &  & X\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a heterogeneous effect of regression (\ref{eq:identity_reg_bias}) on
                      second generation children of parents born in a Spanish speaking country (HH) with different specifications. 
                      I include controls for sex, quartic age, parental education.
                      Standard errors are clustered on the state level}
\item[2] \footnotesize{The samples include second-generation Hispanic children ages 17 and below who live in intact families. 
                      Native born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish speaking 
                      country.}
\item[3] \footnotesize{Data source is the 2004-2020 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}