### 2023-09-02 my_paper/tables/tab50-regression_tab_bygen_skin-90s.tex

\begin{table}[!h]

\caption{Self-Reported Hispanic Identity and Charles and Guryan (2012) Prejeduice Measure: By Generation \label{regtab-bygen-cg-02}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ All Gens \\ $H_i$} & \specialcell{(2) \\  First Gen \\ $H^1_i$} & \specialcell{(3) \\  Second Gen \\ $H^2_i$} & \specialcell{(4) \\  Third Gen \\ $H^3_i$}\\
\midrule
Bias & \num{-0.10}* & \num{-0.03} & \num{-0.09}* & \num{-0.26}**\\
 & (\num{0.05}) & (\num{0.02}) & (\num{0.05}) & (\num{0.12})\\
Female & \num{0.00} & \num{-0.01} & \num{0.01} & \num{-0.03}**\\
 & (\num{0.00}) & (\num{0.01}) & (\num{0.00}) & (\num{0.01})\\
College Graduate: Mother & \num{-0.10}*** & \num{-0.05}** & \num{-0.14}*** & \num{-0.12}***\\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03})\\
College Graduate: Father & \num{-0.13}*** & \num{-0.04} & \num{-0.17}*** & \num{-0.22}***\\
 & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.06})\\
\midrule
N & \num{157424} & \num{21888} & \num{105064} & \num{30472}\\
Mean & \num{0.83} & \num{0.96} & \num{0.87} & \num{0.58}\\
R squared & \num{0.384} & \num{0.018} & \num{0.223} & \num{0.249}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of equation (\ref{eq:identity_reg_bias}) by 
                      parents' type with Charles and Guryan (2012) Prejeduice measure. 
                      Charles and Guryan (2012) use the General Security Survey of the most common racial questions between 1970-2000 for their measure of prejeduice.
                      To use Charles and Guryan (2012) prejeduice measure, I merge the GSS index in 1996 with CPS 1995-1999 years. 
                      In other words, I merge CPS data with the residual prejeduice measure from 20 years before the survey.
                      I include controls for sex, quartic age, parental education.
                      Standard errors are clustered on the state level}
\item[2] \footnotesize{The samples include children ages 17 and below who live in intact families. 
                      First-generation Hispanic immigrant children that were born in a 
                      Spanish speaking county. Native born second-generation Hispanic 
                      immigrant children with at least one parent born in a Spanish speaking 
                      country. Finally, native born third generation Hispanic immigrant children 
                      with native born parents and at least one grand parent born in a Spanish 
                      speaking country.}
\item[3] \footnotesize{Data source is the 2004-2020 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}