### 2023-09-02 my_paper/tables/tab55-regression_tab_bygen_skin-frac-hisp.tex

\begin{table}[!h]
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ All Gens \\ $H_i$} & \specialcell{(2) \\  First Gen \\ $H^1_i$} & \specialcell{(3) \\  Second Gen \\ $H^2_i$} & \specialcell{(4) \\  Third Gen \\ $H^3_i$}\\
\midrule
\cellcolor{yellow}{\textcolor{black}{\textbf{Bias}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{\num{-0.10}**}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{\num{-0.07}*}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{\num{-0.13}**}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{\num{-0.02}}}}\\
\cellcolor{yellow}{\textcolor{black}{\textbf{}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{(\num{0.04})}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{(\num{0.04})}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{(\num{0.05})}}} & \cellcolor{yellow}{\textcolor{black}{\textbf{(\num{0.07})}}}\\
Female & \num{0.00} & \num{0.00} & \num{0.00} & \num{0.00}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
College Graduate: Mother & \num{-0.05}*** & \num{-0.03}*** & \num{-0.06}*** & \num{-0.07}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & \vphantom{1} (\num{0.01})\\
College Graduate: Father & \num{-0.06}*** & \num{-0.06}*** & \num{-0.08}*** & \num{-0.08}***\\
 & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
\midrule
N & \num{844481} & \num{85390} & \num{560100} & \num{198991}\\
Mean & \num{0.91} & \num{0.96} & \num{0.94} & \num{0.82}\\
Year\timesRegion FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\end{table}