### 2023-09-03 output/tables/tab01-SumStat.tex

\begin{table}

\caption{CPS Summary Statistics with Skin IAT Scores \label{tab:sumstat1}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Overall} & \multicolumn{3}{c}{By Generation} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-5}
Characteristic & All Sample 
N = 318,404 & First 
N=40,033 & Second 
N=199,294 & Third 
N=79,077\\
\midrule
Female & 0.49 & 0.54 & 0.49 & 0.49\\
Asian & 0.65 & 0.95 & 0.72 & 0.33\\
Age & 8.7 
(5.1) & 11.2 
(4.5) & 8.5 
(5.1) & 7.7 
(5.0)\\
College Graduate: 
 	 Father & 0.50 & 0.56 & 0.49 & 0.48\\
College Graduate: 
 	 Mother & 0.48 & 0.55 & 0.47 & 0.48\\
Total Family Income 
 	 (1999 dollars) & 78,763 
(80,171) & 69,732 
(72,672) & 78,909 
(81,395) & 82,811 
(80,254)\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\item[1] The samples include children ages 17 and below who live in intact families. First-generation Asian immigrant children that were born in a Asian country. Native-born second-generation Asian immigrant children with at least one parent born in a Asian country. Finally, native-born third generation Asian immigrant children with native-born parents and at least one grand parent born in a Asian country.
\item[2] Data source is the 2004-2021 Current Population Survey.
\end{tablenotes}
\end{threeparttable}
\end{table}