### 2023-09-03 output/tables/tab01a-SumStatOverAll.tex

\begin{table}

\caption{Current Population Survey (CPS) Summary Statistics \label{tab:sumstat1}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{lc}
\toprule
Characteristic & All Sample 
N = 318,404\\
\midrule
Female & 0.49\\
Asian & 0.65\\
Age & 8.7 
(5.1)\\
College Graduate: 
 	 Father & 0.50\\
College Graduate: 
 	 Mother & 0.48\\
Total Family Income 
 	 (1999 dollars) & 78,763 
(80,171)\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The samples include children ages 17 and below who live in intact families. First-generation Asian immigrant children that were born in a Asian country. Native-born second-generation Asian immigrant children with at least one parent born in a Asian country. Finally, native-born third generation Asian immigrant children with native-born parents and at least one grand parent born in a Asian country.
\item[2] Data source is the 2004-2021 Current Population Survey.
\end{tablenotes}
\end{threeparttable}
\end{table}