### 2023-09-05 output/tables/tab05-iat_regression_tab_byparent_type.tex
edit all codes, acs codes might not  be working

\begin{table}[H]

\caption{Relationship Between Bias and Self-Reported Asian identity Among Second-Generation Asian Immigrants: By Parental Type \label{regtab-byparent-01}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{\makecell[c]{Both Parents \\ from Spanish \\ Speaking Country \\ (HH)}} & \multicolumn{1}{c}{\makecell[c]{Father \\ from Spanish \\ Speaking Country \\ (HW)}} & \multicolumn{1}{c}{\makecell[c]{Mother  \\ from Spanish \\ Speaking Country \\ (WH)}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
  & \specialcell{(1) \\ $A^2_{ist}$} & \specialcell{(2) \\ $A^2_{ist}$} & \specialcell{(3) \\ $A^2_{ist}$} & \specialcell{(4) \\ $A^2_{ist}$}\\
\midrule
Bias & -0.45*** & -0.15** & -1.16*** & -0.77***\\
 & (0.09) & (0.07) & (0.23) & (0.20)\\
Female & 0.00 & 0.00 & -0.01 & 0.02\\
 & (0.00) & (0.00) & (0.01) & (0.01)\\
College Graduate: Mother & 0.01 & -0.01 & 0.06* & 0.01\\
 & (0.01) & (0.01) & (0.03) & (0.02)\\
College Graduate: Father & 0.01 & 0.01* & -0.03 & 0.01\\
 & (0.01) & (0.01) & (0.03) & (0.03)\\
\midrule
Observations & 167,709 & 103,697 & 19,982 & 44,030\\
Year $\times$ Region FE & X & X & X & X\\
Mean & 0.72 & 0.96 & 0.38 & 0.32\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a heterogeneous effect of regression (\ref{eq:identity_reg_bias}) by 
                      type of parents with region × year fixed effects. 
                      I include controls for sex, quartic age, fraction of Asians in a state, and parental education.
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include second-generation Asian children ages 17 and below who live in intact families. 
                      Native-born second-generation Asian 
                      immigrant children with at least one parent born in a Asian 
                      country.}
\item[3] \footnotesize{Column (1) includes the results to regression (\ref{eq:identity_reg_bias}) on all second-generation immigrants, 
                                        column (2) includes the results to regression (\ref{eq:identity_reg_bias}) on second-generation immigrants that who has a father and mother that were born in a Asian country (HH),
                                        column (3) includes the results to regression (\ref{eq:identity_reg_bias}) on second-generation immigrants that who has a father that was born in a Asian country and a native-born mother (HW), and
                                        column (4) includes the results to regression (\ref{eq:identity_reg_bias}) on second-generation immigrants that who has a native-born father and a mother that was born in a Asian country (WH).}
\item[4] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}