### 2023-09-05 output/tables/tab45-iat_regression_tab_bygrandparent_type.tex

\begin{table}[H]

\caption{Relationship Between Bias and Self-Reported Asian identity Among Third-Generation Asian Immigrants: By Grandparental Type \label{regtab-bygrandparents}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Number of Asian Grandparents} \\
\cmidrule(l{3pt}r{3pt}){2-5}
  & \specialcell{(1) \\ One} & \specialcell{(2) \\ Two} & \specialcell{(3) \\ Three} & \specialcell{(4) \\ Four}\\
\midrule
Bias & -0.24* & -0.60*** & -2.34*** & -0.24\\
 & (0.14) & (0.17) & (0.78) & (0.20)\\
Female & 0.00 & 0.00 & 0.01 & -0.02*\\
 & (0.01) & (0.02) & (0.03) & (0.01)\\
College Graduate: Mother & 0.01 & 0.04** & 0.11 & -0.02\\
 & (0.01) & (0.02) & (0.09) & (0.02)\\
College Graduate: Father & -0.03*** & 0.00 & -0.03 & 0.02**\\
 & (0.01) & (0.03) & (0.06) & (0.01)\\
\midrule
Observations & 30,052 & 26,255 & 1,324 & 11,751\\
Year $\times$ Region FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of equation (\ref{eq:identity_reg_bias}) restricted to third-generation Asian immigrants by 
                      number of Asian grandparents with region × year fixed effects. 
                      I include controls for sex, quartic age, fraction of Asians in a state, and parental education.
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include third-generation Asian children ages 17 and below who live in intact families. 
                      Native-born third-generation Asian 
                      immigrant children with at least one grandparent born in a Asian 
                      country.}
\item[3] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}