### 2023-09-16 my_paper/tables/tab34-regression_tab_bygen_skin.tex

\begin{table}[!h]

\caption{Subjective Asian Child identity and Asian IAT: By Generation (State) \label{regtab-bygen-04}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ $H_i$} & \specialcell{(2) \\ $H^1_i$} & \specialcell{(3) \\ $H^2_i$} & \specialcell{(4) \\ $H^3_i$}\\
\midrule
Implicit Bias & -0.62*** & -0.14* & -0.54*** & -0.57***\\
 & (0.09) & (0.08) & (0.08) & (0.10)\\
Female & -0.01* & -0.02*** & 0.00 & 0.00\\
 & (0.00) & (0.01) & (0.00) & (0.01)\\
College Graduate: Mother & 0.00 & -0.01 & 0.00 & 0.01\\
 & (0.01) & (0.01) & (0.01) & (0.01)\\
College Graduate: Father & 0.02*** & 0.02** & 0.01 & -0.01\\
 & (0.00) & (0.01) & (0.01) & (0.01)\\
\midrule
N & 268677 & 31586 & 167709 & 69382\\
Mean & 0.65 & 0.95 & 0.72 & 0.33\\
Year-Region FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}