### 2023-09-16 output/tables/maineffect-proxy.tex

\begin{table}[H]

\caption{Main Effect of Proxy on Second-Generation's Asian Self-identification \label{tab:hispbyproxy}}
\centering
\fontsize{12}{14}\selectfont
\begin{tabular}[c]{>{}lllll}
\toprule
Parents Type & All & Asian-Asian & Asian-White & White-Asian\\
\midrule
\textbf{Proxy:} &  &  &  & \\
\hspace{1em}\textbf{Mother} & 0.69 & 0.97 & 0.35 & 0.32\\
\hspace{1em}\textbf{Father} & 0.71 & 0.96 & 0.41 & 0.29\\
\hspace{1em}\textbf{Self} & 0.84 & 0.96 & 0.36 & 0.4\\
\hspace{1em}\textbf{Others} & 0.82 & 0.96 & 0.47 & 0.48\\
\bottomrule
\end{tabular}
\end{table}