### 2025-01-26 output/tables/tab04-regression_tab_bygen_skin.tex

\begin{table}[!h]
\centering\centering
\caption{Relationship Between Bias and Self-Reported Asian Identity: By Generation \label{regtab-bygen-01}}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ $A_i$} & \specialcell{(2) \\ $A^1_i$} & \specialcell{(3) \\ $A^2_i$} & \specialcell{(4) \\ $A^3_i$}\\
\midrule
Bias & -0.09*** & -0.05 & -0.08** & -0.08**\\
 & (0.03) & (0.04) & (0.03) & (0.03)\\
Female & -0.01* & -0.01 & 0.00 & -0.01\\
 & (0.00) & (0.01) & (0.00) & (0.01)\\
College Graduate: Mother & 0.01 & -0.01 & 0.01 & 0.03\\
 & (0.01) & (0.01) & (0.01) & \vphantom{1} (0.02)\\
College Graduate: Father & 0.01** & 0.02** & 0.00 & -0.01\\
 & (0.01) & (0.01) & (0.01) & (0.02)\\
\midrule
Observations & 129,078 & 15,499 & 80,137 & 33,442\\
Mean & 0.65 & 0.44 & 0.22 & 0.66\\
Year $\times$ Region FE & X & X & X & X\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\small
\item[1] \footnotesize{Each column is an estimation of a heterogeneous effect of regression (\ref{eq:identity_reg_bias}) by 
                      generation with region × year fixed effects. 
                      I include controls for sex, quartic age, fraction of Asians in a state, and parental education.
                      I also added parents' (AA, AW, and WA) and grandparents' (AAAA, AAAW, AAWA, etc.) type dummy variables to the regression
                      on second and third generation immigrants, where A is objectively Asian (born in a Asian country) and W is objectively White (native-born). 
                      Standard errors are clustered on the state level.}
\item[2] \footnotesize{The samples include children ages 17 and below who live in intact families. 
                      First-generation Asian immigrant children that were born in a 
                      Asian country. Native-born second-generation Asian 
                      immigrant children with at least one parent born in a Asian 
                      country. Finally, native-born third-generation Asian immigrant children 
                      with native-born parents and at least one grandparent born in a Asian country.}
\item[3] \footnotesize{Data source is the 2004-2021 Current Population Survey.}
\end{tablenotes}
\end{threeparttable}}
\end{table}