### 2025-07-29 submissions/02-jpam/tab01a-SumStatOverAll.tex

\begin{table}
\centering
\caption{CPS Summary Statistics \label{tab:sumstat1}}
\centering
\begin{threeparttable}
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{lc}
\toprule
Characteristic & All Sample N = 318,404\\
\midrule
Female & 0.49\\
Asian & 0.65\\
Age & 8.4 (5.1)\\
College Graduate:  	 Father & 0.52\\
College Graduate:  	 Mother & 0.52\\
Total Family Income  	 (1999 dollars) & 87,031 (84,797)\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\item[1] The sample includes children ages 17 and below who live in intact families. First-generation Asian immigrant children that were born in a Asian country. Native-born second-generation Asian immigrant children with at least one parent born in a Asian country. Finally, native-born third-generation Asian immigrant children with native-born parents and at least one grand parent born in a Asian country.
\item[2] Data source is the 2004-2021 Current Population Survey.
\end{tablenotes}
\end{threeparttable}
\end{table}