### 2025-10-16 submissions/03-jpam-RR/tab58-mnl_third_gen_marginal_effects.tex

\begin{table}[H]
\centering
\caption{Multinomial Logit — Third Generation (AMEs)}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lrrrrrr}
\toprule
variable\_label & est\_ci\_Asian\_only & est\_ci\_White\_only & est\_ci\_Asian\_and\_White & p.value\_Asian\_only & p.value\_White\_only & p.value\_Asian\_and\_White\\
\midrule
Anti-Asian Bias & -0.0764 [-0.0764, -0.0764] & 0.0154 [0.0154, 0.0154] & 0.0610 [0.0610, 0.0610] & 0 & 0 & 0\\
College Graduate: Father & -0.0267 [-0.0267, -0.0267] & 0.0059 [0.0059, 0.0059] & 0.0208 [0.0208, 0.0208] & 0 & 0 & 0\\
College Graduate: Mother & 0.0147 [0.0147, 0.0147] & -0.0428 [-0.0428, -0.0428] & 0.0281 [0.0281, 0.0281] & 0 & 0 & 0\\
Female & 0.0006 [0.0006, 0.0006] & -0.0740 [-0.0740, -0.0740] & 0.0734 [0.0734, 0.0734] & 0 & 0 & 0\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] footnotesize{Average Marginal Effects (AME) with 95% CI.}
\item[2] footnotesize{Standard errors clustered at the state level.}
\item[3] footnotesize{Reference category is embedded in the multinomial baseline (Asian only).}
\item[4] footnotesize{Models include region-year fixed effects.}
\end{tablenotes}
\end{threeparttable}}
\end{table}