### 2025-10-16 output/tables/tab53-SumStatOverAll-adults.tex

\begin{table}

\caption{Current Population Survey (CPS) Summary Statistics \label{tab:sumstat-adults}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{lccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{\textbf{Overall}} & \multicolumn{2}{c}{\textbf{By Generation}} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-4}
\textbf{Characteristic} & \makecell[c]{\textbf{All Sample} \\N = 48,153} & \makecell[c]{\textbf{First} \\N=35,728} & \makecell[c]{\textbf{Second} \\N=12,425}\\
\midrule
Female & 0.56 & 0.58 & 0.50\\
Asian & 0.90 & 0.96 & 0.73\\
Age & 46 (17) & 49 (16) & 38 (17)\\
Years of Education & 14.0 (3.5) & 13.9 (3.8) & 14.5 (2.5)\\
Total Family Income\ \ 	 (1999 dollars) & 68,121 (71,266) & 66,093 (70,444) & 73,950 (73,269)\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The samples include people of Asian ancestry ages 18 and above. First-generation Asian immigrants were born in a Asian country. Native-born second-generation Asian immigrants have at least one parent born in a Asian country. 
\item[2] Data source is the 2004-2021 Current Population Survey.
\end{tablenotes}
\end{threeparttable}
\end{table}