### 2025-10-16 output/tables/tab35-asianiat_regression_tab_byparent_type.tex
WIP: local snapshot before syncing with origin/main

\begin{table}[!h]
\centering\centering
\caption{Subjective Asian Identity and Implicit Bias Among Second Generation Asian Immigrants: By Parental Type (State)\label{regtab-byparent-06}}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{lcccccccccccccccc}
\toprule
\multicolumn{1}{c}{Sample} & \multicolumn{4}{c}{All Sample} & \multicolumn{4}{c}{March Sup.} & \multicolumn{4}{c}{non-March Sup.} & \multicolumn{4}{c}{March Sup. and Control for Income} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13} \cmidrule(l{3pt}r{3pt}){14-17}
\multicolumn{1}{c}{Parents Type} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{AA} & \multicolumn{1}{c}{AW} & \multicolumn{1}{c}{WA} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{AA} & \multicolumn{1}{c}{AW} & \multicolumn{1}{c}{WA} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{AA} & \multicolumn{1}{c}{AW} & \multicolumn{1}{c}{WA} & \multicolumn{1}{c}{All} & \multicolumn{1}{c}{AA} & \multicolumn{1}{c}{AW} & \multicolumn{1}{c}{WA} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9} \cmidrule(l{3pt}r{3pt}){10-10} \cmidrule(l{3pt}r{3pt}){11-11} \cmidrule(l{3pt}r{3pt}){12-12} \cmidrule(l{3pt}r{3pt}){13-13} \cmidrule(l{3pt}r{3pt}){14-14} \cmidrule(l{3pt}r{3pt}){15-15} \cmidrule(l{3pt}r{3pt}){16-16} \cmidrule(l{3pt}r{3pt}){17-17}
  & \specialcell{(1) \\ $H^2$} & \specialcell{(2) \\ $H^2$} & \specialcell{(3) \\ $H^2$} & \specialcell{(4) \\ $H^2$} & \specialcell{(5) \\ $H^2$} & \specialcell{(6) \\ $H^2$} & \specialcell{(7) \\ $H^2$} & \specialcell{(8) \\ $H^2$} & \specialcell{(9) \\ $H^2$} & \specialcell{(10) \\ $H^2$} & \specialcell{(11) \\ $H^2$} & \specialcell{(12) \\ $H^2$} & \specialcell{(13) \\ $H^2$} & \specialcell{(14) \\ $H^2$} & \specialcell{(15) \\ $H^2$} & \specialcell{(16) \\ $H^2$}\\
\midrule
Implicit Bias & -0.09*** & -0.05 & -0.17 & -0.14** & -0.10** & -0.08 & -0.30** & -0.05 & -0.09*** & -0.04 & -0.15 & -0.15** & -0.10** & -0.08* & -0.29** & -0.06\\
 & (0.03) & (0.03) & (0.10) & (0.07) & (0.04) & (0.05) & (0.12) & (0.07) & (0.03) & (0.03) & (0.10) & (0.07) & (0.04) & (0.05) & (0.12) & (0.07)\\
Female & 0.00 & -0.01 & -0.03* & 0.01 & 0.00 & -0.01 & -0.02 & 0.03 & 0.00 & -0.01 & -0.04* & 0.01 & 0.00 & -0.01 & -0.01 & 0.03\\
 & (0.00) & (0.00) & (0.02) & (0.02) & (0.01) & (0.00) & (0.02) & (0.02) & (0.00) & (0.00) & (0.02) & (0.02) & (0.01) & (0.00) & (0.03) & (0.02)\\
College Graduate: Mother & 0.01 & -0.01* & 0.10*** & 0.03 & 0.02 & -0.02* & 0.11** & 0.07** & 0.01 & -0.01* & 0.10*** & 0.02 & 0.02 & -0.02* & 0.12** & 0.07**\\
 & (0.01) & (0.01) & (0.04) & (0.02) & (0.01) & (0.01) & (0.05) & (0.03) & (0.01) & (0.01) & (0.03) & (0.02) & (0.01) & (0.01) & (0.05) & (0.03)\\
College Graduate: Father & 0.00 & 0.01 & -0.06** & 0.00 & 0.00 & 0.02 & -0.08** & -0.03 & 0.00 & 0.01 & -0.06** & 0.01 & 0.00 & 0.02 & -0.07* & -0.02\\
 & (0.01) & (0.01) & (0.02) & (0.03) & (0.01) & (0.01) & (0.04) & (0.04) & (0.01) & (0.01) & (0.02) & (0.03) & (0.01) & (0.01) & (0.04) & (0.04)\\
Log Total Family Income &  &  &  &  &  &  &  &  &  &  &  &  & 0.00 & 0.00** & -0.01 & -0.01\\
 &  &  &  &  &  &  &  &  &  &  &  &  & (0.00) & (0.00) & (0.01) & (0.01)\\
\midrule
Mean & 0.73 & 0.97 & 0.39 & 0.30 & 0.72 & 0.96 & 0.39 & 0.30 & 0.73 & 0.97 & 0.39 & 0.30 & 0.72 & 0.96 & 0.39 & 0.30\\
N & 80137 & 50835 & 9055 & 20247 & 10337 & 6469 & 1253 & 2615 & 69800 & 44366 & 7802 & 17632 & 10337 & 6469 & 1253 & 2615\\
Year-Region FE & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X\\
\bottomrule
\multicolumn{17}{l}{\rule{0pt}{1em}\textsuperscript{1} I include controls for sex, quartic age, parental education.}\\
\multicolumn{17}{l}{\rule{0pt}{1em}\textsuperscript{2} Standard errors are clustered on the state level}\\
\multicolumn{17}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}}
\end{table}